---
title: "Hardcore World of Warcraft Deaths"
author: "Cole Baril"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.asp = .7, out.width = "90%", echo = FALSE, warning = FALSE, message = FALSE)

require(pacman)
p_load(tidyverse, janitor, here, jsonlite, openxlsx, readxl, gt, openxlsx, gtExtras, png, grid, patchwork, randomcoloR, shadowtext, extrafont, ggtext)

options(scipen=999)
```

<style>
mark {
  background-color: #e6cc80; /* Twitter blue background */
  color: black; /* White text color for contrast */
  padding: 0.2em; /* Optional: add some padding for better visual appearance */
  border-radius: 3px; /* Optional: add rounded corners */
}
</style>


<style>
body, p, h1, h2, h3, h4, h5, h6 {
  background-color: white;
  color: black; 
  font-family: Courier New;
}
</style>

```{r, data}
# df1 <- fromJSON(here("hc-wow/Data/db.json"))
# 
# df2 <- fromJSON(here("hc-wow/Data/db2.json"))
# 
# df <- c(df1, df2)

zones <- read_csv(here("hc-wow/Data/zones.csv")) |>  clean_names() 

classes <- read_csv(here("hc-wow/Data/ChrClasses.csv")) |>  select(1, "ID")

creatures <- read_csv(here("hc-wow/Data/creatures.csv")) |>  clean_names()

instances <- read_csv(here("hc-wow/Data/instances.csv"))

# Function to unnest map coordinates
# unnest_element <- function(element) {
#   map_pos <- element$map_pos
#   element$map_pos <- NULL
#   element_df <- as.data.frame(element)
#   element_df$map_pos1 <- map_pos[1]
#   element_df$map_pos2 <- map_pos[2]
#   return(element_df)
# }
# 
# unnested_list <- map(df, unnest_element)
# 
# final_df <- bind_rows(unnested_list, .id = "row_id") |>  
#   left_join(zones, by = c("map_id" = "id")) |>  
#   left_join(classes, by = c("class_id" = "ID")) |>  
#   left_join(creatures, by = c("source_id" = "entry")) |>  
#   left_join(instances, by = c("instance_id" = "instance_id")) |>  
#   rename(player_level = level.x,
#          source_level = level.y,
#          player_class = Name_lang,
#          npc = name)
# 
# final_df <- final_df |> 
#   filter(between(player_level, 1, 60)) |>  
#   rename(damage_source = npc)
# 
# 
# final_df <- final_df |>  
#   mutate(damage_type = case_when(source_id < -1 ~ "Environmental",
#                                  source_id == -1 ~ "PvP",
#                                  TRUE ~ "NPC")) |>  
#   mutate(damage_source = case_when(source_id == -2 ~ "Drowning",
#                                    source_id == -3 ~ "Falling",
#                                    source_id == -4 ~ "Fatigue",
#                                    source_id == -5 ~ "Fire",
#                                    source_id == -6 ~ "Lava",
#                                    source_id == -7 ~ "Slime",
#                                    source_id == -1 ~ "PvP",
#                                    TRUE ~ damage_source))

# final_df |> 
#   write.csv(here("hc-wow/Data/hc-wow_final_df.csv"))

final_df <- read_csv(here("hc-wow/Data/hc-wow_final_df.csv")) |> select(-1)

```

```{r, classic_data}

class_pop <- read_csv(here("hc-wow/Data/vanilla_class_pop.csv")) |>  clean_names() |>  
  summarise(n = sum(parses), .by = class) |>  
  mutate(percent_parses = (n/sum(n))*100) |>  
  select(-n)

```

```{r, colours_icons}
c25 <- c(
  "dodgerblue2", "#E31A1C", 
  "green4",
  "#6A3D9A", 
  "#FF7F00", 
  "black", "gold1",
  "skyblue2", "#FB9A99", 
  "palegreen2",
  "#CAB2D6", 
  "#FDBF6F", 
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

class_colours <- 
  c("Warrior" = "#C79C6E",
    "Paladin" = "#F58CBA",
    "Hunter" = "#ABD473",
    "Rogue" = "#FFF569",
    "Priest" = "black",
    "Shaman" = "#0070DE",
    "Mage" = "#69CCF0",
    "Warlock" = "#9482C9",
    "Druid" = "#FF7D0A")

class_icons <- tibble(class = c("Warrior", "Paladin", "Hunter", "Rogue",
                                "Priest", "Shaman", "Mage", "Warlock", "Druid"),
                      icon.path = c(here("hc-wow/Data/Icons/warrior.png"),
                                    here("hc-wow/Data/Icons/paladin.png"),
                                    here("hc-wow/Data/Icons/hunter.png"),
                                    here("hc-wow/Data/Icons/rogue.png"),
                                    here("hc-wow/Data/Icons/priest.png"),
                                    here("hc-wow/Data/Icons/shaman.png"),
                                    here("hc-wow/Data/Icons/mage.png"),
                                    here("hc-wow/Data/Icons/warlock.png"),
                                    here("hc-wow/Data/Icons/druid.png")))


```

```{r, info_card, fig.width = 15.5, fig.asp = .7, out.width = "90%"}

data <- data.frame(
  x = rep(1:3, times = 2), 
  y = rep(2:1, each = 3),    
  label = c(
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Total Deaths</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(as.integer(nrow(final_df)), big.mark = ",", scientific = FALSE)
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Deaths from NPCs</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(nrow(final_df |> filter(damage_type == "NPC")), big.mark = ","), "<br>", 
      "<span style='color:#14171A; font-size:35px;'>",
      paste0(format(
        (nrow(final_df |> filter(damage_type == "NPC"))) / (nrow(final_df)) * 100, digits = 3
      ), "%")
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Environmental Deaths</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(nrow(final_df |> filter(damage_type == "Environmental")), big.mark = ","), "<br>", 
      "<span style='color:#14171A; font-size:35px;'>",
            paste0(format(
        (nrow(final_df |> filter(damage_type == "Environmental"))) / (nrow(final_df)) * 100, digits = 3
      ), "%")
      
    ),
    
     paste0(
      "<span style='color:#14171A; font-size:30px;'>PvP Deaths</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(nrow(final_df |> filter(damage_type == "PvP")), big.mark = ","), "<br>", 
      "<span style='color:#14171A; font-size:35px;'>",
            paste0(format(
        (nrow(final_df |> filter(damage_type == "PvP"))) / (nrow(final_df)) * 100, digits = 3
      ), "%")
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Unique NPCs</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(length(unique(final_df$damage_source)), big.mark = ",")
    ),
    
     paste0(
      "<span style='color:#14171A; font-size:30px;'>Unique Locations</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(length(unique(final_df$map_name)), big.mark = ",")
      )
    )
  )


ggplot(data) +
  geom_rect(aes(xmin = x - 0.4, xmax = x + 0.4, ymin = y - 0.4, ymax = y + 0.4),
            fill = "#e6cc80", color = "#e6cc80", alpha = 0.5) +
  geom_richtext(aes(x = x, y = y, label = label),
                fill = NA, label.color = NA,  # Transparent background
                size = 5, lineheight = 1.3, hjust = 0.5,
                family = "Courier New", fontface = "bold") +
  scale_x_continuous(expand = c(0, 0), limits = c(0.5, 3.5)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0.5, 2.5)) +
  theme_void(base_family = "Courier New", base_size = 20) +
  labs(title = "Hardcore World of Warcraft Deaths",
       subtitle = "Data from the Deathlog Addon") +
  theme(
    plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 30),
    plot.subtitle = element_text(hjust = 0.5)
  ) 
```

###  <mark> __Number of Deaths__ </mark>

```{r, num_deaths}
final_df |>  
  summarise(n = n(), .by = player_class) |>  
  arrange(desc(n)) |>  
  mutate(percent = (n/sum(n))*100) |>  
  left_join(class_pop, by = c("player_class" = "class")) |> 
  adorn_totals() |>  
  left_join(class_icons, by = c("player_class" = "class")) |>  
  relocate(icon.path, .after = "player_class") |>  
  select(-1) |>  
  rename(player_class = icon.path) |>  
  gt() |>  
  tab_header(title = md("Number of Deaths in Hardcore World of Warcraft")) |>  
  gt_img_rows(columns = player_class, img_source = "local", height = 30) |>  
  cols_label(player_class = md("**Class**"),
             n = md("**No. of Deaths**"),
             percent = md("**% of Total Deaths**"),
             percent_parses = md("**% Vanilla Class Players**")) |>  
  cols_align(align = "center",
             columns = c("n", "percent", "percent_parses")) |>  
  fmt_number(columns = n,
             sep_mark = ",",
             decimals = 0) |>  
  fmt_number(columns = percent:percent_parses,
             decimals = 1) |>  
  opt_table_font(font = "Courier New", weight = "bold") |> 
  tab_options(table.background.color = "#f5f5f2") |> 
  tab_style(
    style = list(
      cell_fill(color = "#C79C6E")
    ),
    locations = cells_body(
      rows = 1
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#ABD473")
    ),
    locations = cells_body(
      rows = 2
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#FFF569")
    ),
    locations = cells_body(
      rows = 3
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#69CCF0")
    ),
    locations = cells_body(
      rows = 4
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#9482C9")
    ),
    locations = cells_body(
      rows = 5
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#FAF9F6")
    ),
    locations = cells_body(
      rows = 6
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#FF7D0A")
    ),
    locations = cells_body(
      rows = 7
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#F58CBA")
    ),
    locations = cells_body(
      rows = 8
    )
  ) |>  
  tab_style(
    style = list(
      cell_fill(color = "#0070DE")
    ),
    locations = cells_body(
      rows = 9
    )
  )
```

###  <mark> __Level Deaths__ </mark> {.tabset}

#### Total Deaths

```{r, level_deaths_total}
level_deaths <- final_df |> 
  group_by(player_level, player_class) |> 
  summarise(n = n()) 

level_deaths |> 
  ggplot(aes(x = player_level, y = n)) +
  geom_point(aes(colour = player_class), size = 2) +
  geom_line(aes(colour = player_class), linewidth = 1) +
  facet_wrap(~ player_class, scales = "free_y") +
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
  scale_x_continuous(breaks = seq(0, 60, 10)) +
  theme(
    strip.text = element_text(face="bold"),
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(size = 18, hjust = 0.5)
  ) +
  scale_colour_manual("Class", values = class_colours) +
  labs(x = "Player Level",
       y = "Number of Deaths",
       title = "Number of Deaths by Level and Class") +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend= -Inf, linewidth = 1) +
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf, linewidth = 1.5) +
  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none",
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank()
        
  )
  
```

#### Normalized Deaths

```{r, level_deaths_normalized}
total_deaths <- level_deaths |> 
  group_by(player_class) |> 
  summarise(total_deaths = sum(n))

# Find the level with the highest normalized deaths for each class
peak_levels <- final_df |> 
  group_by(player_level, player_class) |> 
  summarise(n = n()) |> 
  left_join(total_deaths, by = "player_class") |> 
  mutate(normalized_deaths = n / total_deaths) |>  
  group_by(player_class) |> 
  filter(normalized_deaths == max(normalized_deaths))



total_deaths_all <- level_deaths |> 
  summarise(total_deaths = sum(n))

# Merge the total deaths with the original dataset
final_df <- final_df |> 
  left_join(total_deaths, by = "player_class")

# Create the plot in logspace with normalized deaths


final_df |> 
  group_by(player_level, player_class) |> 
  summarise(n = n()) |> 
  left_join(total_deaths, by = "player_class") |> 
  mutate(normalized_deaths = n / total_deaths) |> 
  ggplot(aes(x = player_level, y = normalized_deaths)) +
  geom_point(aes(colour = player_class), size = 1.5) +
  geom_line(aes(colour = player_class), linewidth = 1) +
  annotate("segment", x = -Inf, xend = Inf, y = 0, yend= 0, linewidth = 1) +
  annotate("segment", x = -Inf, xend = -Inf, y = 0, yend = Inf, linewidth = 1.5) +
  facet_wrap(~ player_class) +
  scale_y_log10(labels = scales::percent) +
  scale_x_continuous(breaks = seq(0, 60, 10)) +
  scale_colour_manual("Class", values = class_colours) +
  labs(x = "Player Level",
       y = "Percent of Total Deaths (Log10 Scale)",
       title = "Percent of Deaths by Level and Class") +
  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none",
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank()
        
  )
```

###  <mark> __Death Source Level v. Player Level__ </mark>

```{r, source_level}
final_df %>% 
  filter(damage_type == "NPC") %>% 
  ggplot(aes(x = player_level, y = source_level)) + 
  geom_smooth(se = FALSE, aes(colour = player_class)) +
  facet_wrap(~player_class) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend= -Inf, linewidth = 1) +
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf, linewidth = 1.5) +
  scale_colour_manual("Class", values = class_colours) +
  theme_bw(base_size = 20, base_family = "Courier New") +
  labs(title = "Death Source Level v. Player Level",
       x = "Player Level",
       y = "Death Source Level") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank())
```

###  <mark> __Deadliest Zones__ </mark>

```{r, zone_deaths}


deadliest_zones <- final_df %>%
  drop_na(map_name) %>% 
  group_by(player_level, map_name) %>%
  summarise(death_count = n()) %>%
  arrange(desc(death_count)) %>% 
  group_by(player_level) %>%
  slice(1) 

deadliest_zones %>% 
  ggplot(aes(x = player_level, y = death_count, fill = map_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual("Zone", values = c25) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(from = 0, to = 60, by = 2)) +
  scale_y_continuous(expand = c(0, 0), labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
  labs(x = "Player Level",
       y = "Number of Deaths",
       title = "Deadliest Zone for Each Level") +
  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(1, "lines"),
        legend.key.height = unit(0.5, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        legend.text = element_text(size = 10),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black")) +
  guides(fill = guide_legend(ncol = 6))
  

```

###  <mark> __Deadliest Instances__ </mark>

```{r, instances}


deadliest_instance <- final_df %>%
  drop_na(instance_name) %>% 
  group_by(player_level, instance_name) %>%
  summarise(death_count = n()) %>%
  arrange(desc(death_count)) %>% 
  group_by(player_level) %>%
  slice(1) 
deadliest_instance %>% 
  ggplot(aes(x = player_level, y = death_count, fill = instance_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual("Instance", values = c25) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(from = 0, to = 60, by = 2)) +
  scale_y_continuous(expand = c(0, 0), labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
  theme(legend.position = "bottom") +
  labs(x = "Player Level",
       y = "Number of Deaths",
       title = "Deadliest Instance for Each Level") +
    theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(1, "lines"),
        legend.key.height = unit(0.5, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        legend.text = element_text(size = 10),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black")) +
  guides(fill = guide_legend(ncol = 6))
```

###  <mark> __Deadliest NPCs__ </mark>

```{r, npc, fig.width = 15}


deadliest_npc <- final_df %>%
  filter(damage_type == "NPC") %>% 
  drop_na(damage_source) %>% 
  group_by(player_level, damage_source) %>%
  summarise(death_count = n()) %>%
  arrange(desc(death_count)) %>% 
  group_by(player_level) %>%
  slice(1) 

n = 37
pal <- distinctColorPalette(n)

npc_levels <- final_df %>% 
  filter(damage_type == "NPC") %>% 
  drop_na(damage_source) %>% 
  distinct(damage_source, .keep_all = TRUE) %>% 
  select(damage_source, source_level)

deadliest_npc %>% 
  left_join(npc_levels, by = "damage_source") %>% 
  ggplot(aes(x = player_level, y = death_count, fill = damage_source)) +
  geom_bar(stat = "identity") +
  geom_shadowtext(aes(label = source_level),
                  hjust = 0.5,
                  vjust = -0.5,
                  position = position_dodge()) +
  scale_fill_manual("NPC", values = pal) +
 scale_x_continuous(expand = c(0, 0), breaks = seq(from = 0, to = 60, by = 2)) +
  scale_y_continuous(expand = c(0, 0), labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
  expand_limits(y = c(0, 4250)) +
  labs(x = "Player Level",
       y = "Number of Deaths",
       title = "Deadliest NPC for Each Level",
       subtitle = "NPC Level Shown Above Bars") +
      theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(0.5, "lines"),
        legend.key.height = unit(0.25, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        legend.text = element_text(size = 8),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black")) +
  guides(fill = guide_legend(ncol = 7))
 
```