---
title: "Twitter: The End of an Era"
author: "Cole Baril"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.asp = .7, out.width = "90%", echo = FALSE, warning = FALSE, message = FALSE)

require(pacman)
p_load(tidyverse, here, jsonlite, data.table, gt, patchwork, hms, tidytext, lubridate, ggtext, reshape2, wordcloud, treemapify, extrafont)
```

<style>
mark {
  background-color: #1DA1F2; /* Twitter blue background */
  color: black; /* White text color for contrast */
  padding: 0.2em; /* Optional: add some padding for better visual appearance */
  border-radius: 3px; /* Optional: add rounded corners */
}
</style>


<style>
body, p, h1, h2, h3, h4, h5, h6 {
  background-color: white;
  color: black; 
  font-family: Courier New;
}
</style>



```{r, import_clean}
# twitter <- fromJSON(here("twitter/Data/tweet.json"), flatten = TRUE) |> 
#   mutate(across(.cols = c(tweet.favorite_count, tweet.retweet_count), ~ as.numeric(.))) |> 
#   mutate(statusreply = case_when(is.na(tweet.in_reply_to_status_id) ~ "Yes",
#                              TRUE ~ "No")) |> 
#   mutate(datetime = parse_date_time(tweet.created_at, "a b d H:M:S z Y", tz = "UTC"),
#          datetime = with_tz(datetime, "America/Chicago"),
#          date = date(datetime),
#          month = month(date),
#          year = year(date),
#          time = as_hms(datetime),
#          weekday = lubridate::wday(date, label = TRUE),
#          hour_floored = as_hms(floor_date(datetime, "hour"))) |> 
#   unnest_wider(col = tweet.display_text_range, names_sep = "_") 

twitter <- fromJSON(here("twitter/Data/new_tweet1.js"), flatten = TRUE) |> 
   mutate(across(.cols = c(tweet.favorite_count, tweet.retweet_count), ~ as.numeric(.))) |> 
   mutate(statusreply = case_when(is.na(tweet.in_reply_to_status_id) ~ "Yes",
                             TRUE ~ "No")) |> 
  mutate(datetime = parse_date_time(tweet.created_at, "a b d H:M:S z Y", tz = "UTC"),
         datetime = with_tz(datetime, "America/Chicago"),
         date = date(datetime),
         month = month(date),
         year = year(date),
         time = as_hms(datetime),
         weekday = lubridate::wday(date, label = TRUE),
         hour_floored = as_hms(floor_date(datetime, "hour"))) |> 
  unnest_wider(col = tweet.display_text_range, names_sep = "_") 
```

```{r, info_card, fig.width = 15.5, fig.asp = .7, out.width = "90%"}





data <- data.frame(
  x = rep(1:3, times = 2), 
  y = rep(2:1, each = 3),    
  label = c(
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Total Tweets</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(as.integer(nrow(twitter)), big.mark = ",", scientific = FALSE)
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Days Spent Tweeting</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", format(length(unique(twitter$date)), big.mark = ",")
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Characters Tweeted</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", 
      format(sum(as.numeric(twitter$tweet.display_text_range_2)), big.mark = ",")
    ),
    
     paste0(
      "<span style='color:#14171A; font-size:30px;'>Interacted With</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", 
      format(length(unique(twitter$tweet.in_reply_to_status_id_str)), big.mark = ","),
      "<span style='color:#14171A; font-size:35px;'>", " accounts"
    ),
    
    paste0(
      "<span style='color:#14171A; font-size:30px;'>Received</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", 
      format(sum(as.numeric(twitter$tweet.favorite_count)), big.mark = ","),
       "<span style='color:#14171A; font-size:35px;'>", 
       " favourites"
    ),
    
     paste0(
      "<span style='color:#14171A; font-size:30px;'>Received</span><br><br>",
      "<span style='color:#14171A; font-size:40px;'>", 
      format(sum(as.numeric(twitter$tweet.retweet_count)), big.mark = ","),
       "<span style='color:#14171A; font-size:35px;'>", 
       " retweets"
      )
    )
  )


ggplot(data) +
  geom_rect(aes(xmin = x - 0.4, xmax = x + 0.4, ymin = y - 0.4, ymax = y + 0.4),
            fill = "#1DA1F2", color = "#14171A", alpha = 0.5) +
  geom_richtext(aes(x = x, y = y, label = label),
                fill = NA, label.color = NA,  # Transparent background
                size = 5, lineheight = 1.3, hjust = 0.5,
                family = "Courier New", fontface = "bold") +
  scale_x_continuous(expand = c(0, 0), limits = c(0.5, 3.5)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0.5, 2.5)) +
  theme_void(base_family = "Courier New", base_size = 20) +
  labs(title = "Twitter: The End of an Era",
       subtitle = "2015 to 2024: 9 Years of Tweets Until I Left for Bluesky") +
  theme(
    plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 30),
    plot.subtitle = element_text(hjust = 0.5)
  ) 
```

### <mark> __Histogram__ </mark>

```{r}
twitter %>% 
  summarize(n = n(),
            .by = date) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(fill = "#1DA1F2", colour = "black") +
  labs(title = "Tweet Histogram",
       x = "Number of Tweets",
       y = "Number of Days Tweeted") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
   theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(2.5, "lines"),
        legend.key.height = unit(1, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))
```

###  <mark> __Twitter Activity Over Time__ </mark> {.tabset}

#### Date by Time

```{r}
twitter |> 
  ggplot(aes(x = date, y = time)) +
  geom_point(colour = "#1DA1F2", size = 1) +
  geom_vline(
    xintercept = date("2022-04-14"), color = "red",  
    linetype = "dashed", size = 1) +
   annotate("text", 
           x = date("2022-04-01"), 
           y = hm("12:00"), 
           label = 'atop(bold("Musk Buys Twitter"))',
           colour = "red",
           parse = TRUE,
           angle = 90,
           size = 5
           ) +
  
   geom_vline(
    xintercept = date("2024-11-15"), color = "black",  
    linetype = "dashed", size = 1) +
   annotate("text", 
           x = date("2024-11-01"), 
           y = hm("12:00"), 
           label = 'atop(bold("I joined Bluesky"))',
           colour = "black",

           parse = TRUE,
           angle = 90,
           size = 5
           ) +
  
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  scale_y_time(breaks = scales::breaks_width("4 hours"), 
               labels = function(l) strftime(l, '%H:%M'),
              
               expand = c(0, 0)) +
  labs(title = "Tweets by Me",
       subtitle = "1 dot = 1 Tweet",
       x = "Date",
       y = "Local Time") +

   theme_bw(base_size = 20, base_family = "Courier New") +
   theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(2.5, "lines"),
        legend.key.height = unit(1, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))

ggsave("tweets_by_me.png", plot = last_plot(), width = 12,, height = 9, dpi = 500)
  
```

#### Year by Hour

```{r}
twitter %>% 
  mutate(hour_floor = floor_date(datetime, "hour")) %>%
  mutate(hour = as_hms(hour_floor)) %>% 
  summarize(n = n(),
            .by = c(year, hour)) %>% 
  mutate(hour = hour(hour)) %>% 
  ggplot(aes(x = hour, y = n)) + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    geom_bar(stat = "identity", position = "dodge", fill = "#1DA1F2", colour = "black") +
    scale_x_continuous(breaks = seq(0, 23, 4)) +
    labs(title = "Number of Tweets Sent by Year and Hour",
        x = "Local Time",
        y = "Number of Tweets") +
    facet_wrap(~year) +
    theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(2.5, "lines"),
        legend.key.height = unit(1, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))
```

#### Year by Weekday

```{r}
twitter %>% 
  summarize(n = n(),
            .by = c(year, weekday)) %>% 
  ggplot(aes(x = weekday, y = n)) + 
  geom_col(fill = "#1DA1F2", colour = "black") +
  labs(title = "Number of Tweets Sent during Days of the Week",
       x = "Day of the Week",
       y = "Number of Tweets") +
 
  facet_wrap(~year) +

  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(2.5, "lines"),
        legend.key.height = unit(1, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))
```

###  <mark> __Tweet Heatmap__ </mark>

```{r, heatmap, fig.width = 10}
twitter |> 
  summarize(n_tweets = n(),
            .by = c(hour_floored, weekday)) |> 
  ggplot(aes(x = weekday, y = hour_floored, fill = n_tweets)) +
  geom_tile(colour = "white", linewidth = 0.1) + 
  scale_fill_gradient(high = "#155A88", low = "#A6DFFC", labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
   scale_y_time(breaks = scales::breaks_width("2 hours"), 
               labels = function(l) strftime(l, '%H:%M'),
               expand = c(0, 0)) +
  theme_bw(base_size = 20) +
  labs(x = "Weekday",
       y = "Local Time",
       fill = "Tweets",
       title = "Twitter Weekly Post Heatmap") +
  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(hjust = 0.5),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank()
  )
```


###  <mark> __Twitter Activity: Favourites & Retweets__ </mark> {.tabset}

#### Favourites

```{r}
twitter |> 
  ggplot(aes(x = date, y = time)) +
  geom_point(aes(color = statusreply, size = tweet.favorite_count), ) +
    scale_colour_manual("Status or Reply?",
    labels = c("Reply", "Status"),
    values = c("black", "#1DA1F2")) +
  
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  scale_y_time(breaks = scales::breaks_width("4 hours"), 
               labels = function(l) strftime(l, '%H:%M')
               # expand = c(0, 0)
               ) +
  scale_size("Number of Favourites", range = c(1, 15)) +
  labs(title = "Tweets by Me: Number of Favourites",
       subtitle = "1 dot = 1 Tweet",
       x = "Date",
       y = "Local Time") +

   theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.size = unit(3, "point"),

        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black")) +
  guides(size = guide_legend(nrow = 1),
         colour = guide_legend(override.aes = list(size = 10)))
```

#### Retweets

```{r}
twitter |> 
  ggplot(aes(x = date, y = time)) +
  geom_point(aes(color = statusreply, size = tweet.retweet_count), ) +
    scale_colour_manual("Status or Reply?",
    labels = c("Reply", "Status"),
    values = c("black", "#1DA1F2")) +
  
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  scale_y_time(breaks = scales::breaks_width("4 hours"), 
               labels = function(l) strftime(l, '%H:%M')
               # expand = c(0, 0)
               ) +
  scale_size("Number of Retweets", range = c(1, 15)) +
  labs(title = "Tweets by Me: Number of Retweets",
       subtitle = "1 dot = 1 Tweet",
       x = "Date",
       y = "Local Time") +

   theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.size = unit(3, "point"),

        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black")) +
  guides(size = guide_legend(nrow = 1),
         colour = guide_legend(override.aes = list(size = 10)))
```


###  <mark> __Tweet Length__ </mark>

```{r}
twitter |> 
  ggplot(aes(x = date, y = as.numeric(tweet.display_text_range_2))) +
  geom_point(colour = "#1DA1F2",
             size = 1) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  geom_vline(xintercept = date("2017-11-07"), linetype = "dashed", colour = "red") +
  geom_hline(yintercept = 140, linetype = "dashed", colour = "red") +
  geom_hline(yintercept = 280, linetype = "dashed", colour = "red") +
  annotate("text", 
           x = date("2017-10-15"), 
           y = 300, 
           label = 'atop(bold("Twitter announces character limit increase to 280"))',
           parse = TRUE,
           angle = 90) +
  annotate("text", 
           x = date("2015-10-01"),
           y = 140,
           label = 'atop(bold("140 Character Limit"))', 
           parse = TRUE) +
  annotate("text", 
           x = date("2015-10-01"),
           y = 280,
           label = 'atop(bold("280 Character Limit"))',
           parse = TRUE) +
  labs(
    title = "Number of Characters in Each Tweet",
    x = "Date",
    y = "Length of Tweet") +
   theme_bw(base_size = 20, base_family = "Courier New") +
   theme(panel.grid.major.x = element_blank(),
        # panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "top",
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5),
        legend.key.width = unit(2.5, "lines"),
        legend.key.height = unit(1, "lines"),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))
```

###  <mark> __Sentiment Frequency Analysis__ </mark>

```{r}
unclean_tweet <- twitter$tweet.full_text

clean_tweets <- function(x) {
  x %>%
    # Remove URLs
    str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
    # Remove mentions e.g. "@my_account"
    str_remove_all("@[[:alnum:]_]{4,}") %>%
    # Remove hashtags
    str_remove_all("#[[:alnum:]_]+") %>%
    # Replace "&" character reference with "and"
    str_replace_all("&amp;", "and") %>%
    # Remove puntucation, using a standard character class
    str_remove_all("[[:punct:]]") %>%
    # Remove "RT: " from beginning of retweets
    str_remove_all("^RT:? ") %>%
    # Replace any newline characters with a space
    str_replace_all("\\\n", " ") %>%
    # Make everything lowercase
    str_to_lower() %>%
    # Remove any trailing whitespace around the text
    str_trim("both")
}

# Pass the clean_tweets function through tweets and convert to a tibble, retaining line (tweet) data.
cleaned_tweets <- unclean_tweet %>% 
  clean_tweets() %>% 
  tibble(line = 1:nrow(twitter)) %>% 
  rename_at(1, ~'text')

# Extract words
word_df <- cleaned_tweets %>% 
  unnest_tokens(word, text)

wordcount <- word_df %>% 
  count(word)


bing <- get_sentiments("bing")

word_df %>%  
  inner_join(get_sentiments("bing")) %>% 
  count(word, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative) %>% 
  ggplot(aes(x = word, y = sentiment)) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  labs(title = "Twitter Word Sentiment: Positive or Negative?",
       subtitle = "The number of times each positive (e.g., > 0) and \nnegative (e.g., < 0) word was Tweeted.",
       y = "Word Frequency",
       x = "Word") +
  
  theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(hjust = 0.5),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line.y = element_line(colour = "black")
        )
```

<!-- ###  <mark> __Sentiment Word Cloud__ </mark> -->

```{r, include = FALSE, results = FALSE}
word_df %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(word, sentiment) %>% 
  acast(word ~ sentiment, value.var = "n", fill = 0) %>% 
  comparison.cloud(colors = c("firebrick4", "forestgreen"),
                   max.words = 100)

```

###  <mark> __NRC Sentiment Tree Map__ </mark>

```{r}
nrc <- get_sentiments("nrc")

word_df %>% 
  inner_join(get_sentiments("nrc")) %>% 
  count(word, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  select_if(is.numeric) %>% 
  map_dbl(sum) %>% 
  as.data.frame() %>% 
  rownames_to_column("Emotion") %>% 
  rename("number" = ".") %>% 
  mutate(across(where(is.character), tools::toTitleCase)) %>% 
  ggplot(aes(area = number, fill = Emotion, label = Emotion)) +
  geom_treemap() + 
  geom_treemap_text(colour = "Black", 
                    place = "centre",
                    size = 25) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Emotion Frequency Based on Tweeted Words") +
   theme_bw(base_size = 20, base_family = "Courier New") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(hjust = 0.5),
        legend.background = element_rect(fill = "#f5f5f2"),
        plot.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.background = element_rect(fill = "#f5f5f2", color = "#f5f5f2"),
        panel.border = element_blank(),
        axis.line.y = element_line(colour = "black")
        )
```